---
title: "Inflation"
description: |
  Will US inflation last another year or more?
author:
  - name: Arham Choudhury
    url: https://arhamchoudhury.com/
date: 2022-08-01
output:
  distill::distill_article:
    self_contained: false
    code_folding: hide
---

*Executive Summary*: The recent interest rate hikes appear to have convinced forecasters and market participants that inflation 
will be tamed to below 4% level within a year. Businesses and individuals should see this as a sign of some relief.

```{r chart echo=FALSE}
chart
```

Note: Dotted line represents expected inflation 1-year from now.
Solid line shows actual reported CPI inflation.
Sources: Actual Inflation: Bureau of Labor Statistics, data is Monthly, not seasonally adjusted
         Expected Inflation: Federal Reserve Bank of Cleveland



```{r code}

library(tidyverse)
suppressPackageStartupMessages(library(tidyquant))
library(timetk)
library(scales)
library(ggrepel)

# Change vs 12 months ago Function
YY <- function(x){
  (x - lag(x,12))/lag(x,12)
}

# Actual Inflation
# As of Aug 2022

CPIAUCNS <- tq_get("CPIAUCNS", get = "economic.data", from = "2018-01-01") %>%
  summarise_by_time(
    .date_var = date,
    .by = "month",
    CPI_allitems_monthly = mean(price, na.rm = T)) %>%
  mutate(CPI_YY = YY(CPI_allitems_monthly)*100) %>%
  na.omit() %>%
  filter(date >= "2022-01-01") %>%
  select(date, CPI_YY) %>%
  rename(value = CPI_YY) %>%
  mutate(type = "Actual")
CPIAUCNS


EXPINF1YR <- tq_get("EXPINF1YR", get = "economic.data", from = "2018-01-01") %>%
  summarise_by_time(
    .date_var = date,
    .by = "month",
    CPI_exp_allitems_monthly = mean(price, na.rm = T)) %>%
  na.omit() %>%
  slice_tail() %>%
  mutate(date = date + months(12)) %>%
  rename(value = CPI_exp_allitems_monthly) %>%
  mutate(type = "Expected")
EXPINF1YR



CPI_actual_plus_exp <- bind_rows(CPIAUCNS, EXPINF1YR) %>%
  mutate(value = round(value,1))
CPI_actual_plus_exp


chart <- ggplot(data = filter(CPI_actual_plus_exp, type == "Actual"),
         aes(x = date, y = value, label = value)) +
  geom_line() +
  geom_point() +
  geom_line(data = filter(CPI_actual_plus_exp, date >= "2022-07-01"),
            aes(x = date, y = value), linetype = "dashed") +
  geom_point(data = filter(CPI_actual_plus_exp, type == "Expected"),
             aes(x = date, y = value)) +
  geom_hline(yintercept = 0) +
  scale_y_continuous(labels = scales::number_format(accuracy = 0.1),
                     breaks = pretty_breaks(10)) +
  scale_x_date(date_breaks = "1 month",
               date_labels = "%m-%y") +
  labs(title = "Inflation: Actual & Expected",
       y     = "Inflation (YoY %)", x = "Month") +
  theme(axis.text.x=element_text(angle=45, hjust=1, size = 8)) +
  geom_text_repel(data = filter(CPI_actual_plus_exp, date >= "2022-07-01"))
chart



```






# end
